Computable Poincar\'e--Friedrichs constants for the $L^{p}$ de Rham complex over convex domains and domains with shellable triangulations

### 2024-10-06 main.tex
more clean up

We wish to clarify the relationship between the Poincar\'e--Friedrichs inequality, in the sense introduced above, with other inequalities that are known as Poincar\'e inequality (or also Poincar\'e--Wirtinger or Friedrichs inequality) in the literature~\cite[Remark~\mwl{todo}]{ern2021finite}. Given $p \in [1,\infty]$ and a domain $\Omega \subseteq \bbR^{n}$ of finite measure, 

@@

% \cye{Recall the notations~\eqref{equation:W1p} and~\eqref{equation:W_curl_div_p} from Section~\ref{section:intro}; actually, here we temporarily rather employ 
% \begin{gather*}
    % W^{p}(\grad,\Omega) \cye{= W^{1,p}(\Omega)} = \{ u \in L^{p}(\Omega) \suchthat \grad v \in \bfL^{p}(\Omega) \}
    % .
% \end{gather*}}
Let $\Omega \subseteq \bbR^{3}$ be a bounded open set. 
We recall $L^{p}(\Omega)$, the space of scalar-valued $p$-integrable functions defined on $\Omega$
and that $\bfL^{p}(\Omega) := L^{p}(\Omega)^{n}$ for vector-valued functions with each component in $L^{p}(\Omega)$. 
In the three-dimensional setting, we are particularly interested in the Sobolev vector analysis. 
The space of scalar-valued $L^{p}(\Omega)$ functions with weak gradients in $\bfL^{p}(\Omega)$ is 
\begin{gather*}
    W^{p}(\grad,\Omega) := W^{1,p}(\Omega) = \{ u \in L^{p}(\Omega) \suchthat \grad v \in \bfL^{p}(\Omega) \}
    .
\end{gather*}
The space $\bfW^{p}(\curl,\Omega)$ of vector-valued $\bfL^{p}(\Omega)$ functions with weak curls in $\bfL^{p}(\Omega)$
and the space of vector-valued $\bfL^{p}(\Omega)$ functions with weak divergences in $L^{p}(\Omega)$ are written 
\begin{gather*}
    \bfW^{p}(\curl,\Omega) = \{ \bfu \in \bfL^{p}(\Omega) \suchthat \curl \bfu \in \bfL^{p}(\Omega) \},
    \\ 
    \bfW^{p}(\divergence,\Omega) = \{ \bfu \in \bfL^{p}(\Omega) \suchthat \divergence \bfu \in L^{p}(\Omega) \}.
\end{gather*}
We are interested in transformations of these Sobolev tensor fields from one domain onto another. 
Suppose that $\Omega, \Omega' \subset \mathbb{R}^3$ are open sets and suppose that $\phi: \Omega \to \Omega'$ is a bi-Lipschitz mapping.
\mwl{We introduce the} gradient-, curl-, and divergence-conforming Piola transformations, respectively, as the mappings 
$\phi^{\grad}: L^{p}(\Omega') \to L^{p}(\Omega)$,
$\phi^{\curl}: \bfL^{p}(\Omega') \to \bfL^{p}(\Omega)$, and 
$\phi^{\divergence}: \bfL^{p}(\Omega') \to \bfL^{p}(\Omega)$.
We also introduce 
$\phi^{\rm b}: L^{p}(\Omega') \to L^{p}(\Omega)$. 
We define them 
for any $v \in L^{p}(\Omega')$ and $\bfw \in \bfL^{p}(\Omega')$ by setting 
\begin{subequations}\label{eq_definition_piola}
\begin{align*}
    \phi^{\grad}(v) &= v \circ \phi, \\
    \phi^{\curl}(\bfw) &= \Jacobian\phi^{T} (\bfw \circ \phi), \\
    \phi^{\divergence}(\bfw) &= \adj(\Jacobian\phi) \left( \bfw \circ \phi \right), \\  
    \phi^{\rm b}(v) &= \det(\Jacobian\phi) \left( v \circ \phi \right),
\end{align*}
\end{subequations}
Here, $\Jacobian\phi$ is the Jacobian matrix of $\phi$ (see also~\cite[Definition~9.8]{ern2021finite}), and $\adj\Jacobian\phi$ denotes taking its adjugate matrix.
These transformations are invertible. 
Bounds on the Lebesgue norms will follow from a more general result below. 
We use the commutativity relations 
\begin{subequations}\label{eq_piola_commute}
\begin{align}
    \grad \phi^{\grad} (v) &= \phi^{\curl}(\grad v), 
    \\
    \curl \phi^{\curl}(\bfv) &= \phi^{\divergence}(\curl \bfv), 
    \\
    \divergence \phi^{\divergence}(\bfw) &= \phi^{\rm b}(\divergence \bfw),
\end{align}
\end{subequations}
where $v \in W^{p}(\grad,\Omega)$, $\bfv \in \bfW^{p}(\curl,\Omega)$, and $\bfw \in \bfW^{p}(\divergence,\Omega)$.  
We summarize this as a commuting diagram:
\begin{align*}
    \begin{CD}
        W^{p}(\grad,\Omega') @>{\grad}>> \bfW^{p}(\curl,\Omega') @>{\curl}>> \bfW^{p}(\divergence,\Omega') @>{\divergence}>> L^{p}(\Omega')
        \\
        @VV{\phi^{\grad}}V 
        @VV{\phi^{\curl}}V 
        @VV{\phi^{\divergence}}V 
        @VV{\phi^{\rm b}}V 
        \\
        W^{p}(\grad,\Omega ) @>{\grad}>> \bfW^{p}(\curl,\Omega ) @>{\curl}>> \bfW^{p}(\divergence,\Omega ) @>{\divergence}>> L^{p}(\Omega ).
    \end{CD}
\end{align*}

@@

\mwl{We now move the discussion to exterior calculus, beginning with exterior algebra.}
Let $V$ be a real vector space. 
Given an integer $k \geq 0$, we let $\Alt^{k}(V)$ denote the space of scalar-valued antisymmetric $k$-linear forms over $V$. 
Recall that any $k$-linear scalar-valued form $u$ over $V$ is called antisymmetric
if 
\begin{gather*} 
    u( v_{\pi(1)}, v_{\pi(2)}, \ldots, v_{\pi(k)} ) 
    = 
    \text{sign}(\pi) 
    u( v_1, v_2, \ldots, v_k ) 
\end{gather*}
for any $v_1, v_2, \dots, v_k \in V$ and any permutation $\pi$ of the indices \(\{1, 2, \ldots, k\}\). 
By definition, $\Alt^{1}(V)$ is just the dual space of $V$, and $\Alt^{0}(V)$ is the space of real numbers. 
Formally, we define $\Alt^{k}(V)$ to be the zero vector space when $k < 0$.

@@

\begin{example}\label{example:shell_simpl}
    Any simplex $T$ (trivially) has a shelling, consisting only of $T$ itself. 
    The boundary complex $\partial \calT(T)$ has a shelling:
    any enumeration of the boundary faces of $T$ constitutes a shelling\mwl{; see Example 8.2.(iii) in~\cite{ziegler1995lectures}}.
\end{example}
\begin{example}
    The standard triangulation of the $3$-dimensional cube by six tetrahedra, the Kuhn triangulation~\cite{kuhn1960some}, is shellable, as are its higher-dimensional generalizations.\footnote{We remark that Kuhn attributes this triangulation to Lefschetz~\cite{lefschetz2015introduction}.}
\end{example}
\begin{example}
    There exists a non-shellable triangulation of a tetrahedron and of a cube in $n=3$, see~\cite[Example 8.9]{ziegler1995lectures}.
\end{example}

@@

        % TODO Check that $\calS'$ is a triangulation...

>>>>

We wish to clarify the relationship between the Poincar\'e--Friedrichs inequality, in the sense introduced above, with other inequalities that are known as Poincar\'e inequality (or also Poincar\'e--Wirtinger or Friedrichs inequality) in the literature~\cite[Remark~3.32]{ern2021finite}. Given $p \in [1,\infty]$ and a domain $\Omega \subseteq \bbR^{n}$ of finite measure, 

@@

% \cye{Recall the notations~\eqref{equation:W1p} and~\eqref{equation:W_curl_div_p} from Section~\ref{section:intro}; actually, here we temporarily rather employ 
% \begin{gather*}
    % W^{p}(\grad,\Omega) \cye{= W^{1,p}(\Omega)} = \{ u \in L^{p}(\Omega) \suchthat \grad v \in \bfL^{p}(\Omega) \}
    % .
% \end{gather*}}
Let $\Omega \subseteq \bbR^{3}$ be a bounded open set. 
We recall $L^{p}(\Omega)$, the space of scalar-valued $p$-integrable functions defined on $\Omega$
and that $\bfL^{p}(\Omega) := L^{p}(\Omega)^{n}$ for vector-valued functions with each component in $L^{p}(\Omega)$. 
In the three-dimensional setting, we are particularly interested in the Sobolev vector analysis. 
The space of scalar-valued $L^{p}(\Omega)$ functions with weak gradients in $\bfL^{p}(\Omega)$ is 
\begin{gather*}
    W^{p}(\grad,\Omega) := W^{1,p}(\Omega) = \{ u \in L^{p}(\Omega) \suchthat \grad v \in \bfL^{p}(\Omega) \}
    .
\end{gather*}
The space $\bfW^{p}(\curl,\Omega)$ of vector-valued $\bfL^{p}(\Omega)$ functions with weak curls in $\bfL^{p}(\Omega)$
and the space of vector-valued $\bfL^{p}(\Omega)$ functions with weak divergences in $L^{p}(\Omega)$ are written 
\begin{gather*}
    \bfW^{p}(\curl,\Omega) = \{ \bfu \in \bfL^{p}(\Omega) \suchthat \curl \bfu \in \bfL^{p}(\Omega) \},
    \\ 
    \bfW^{p}(\divergence,\Omega) = \{ \bfu \in \bfL^{p}(\Omega) \suchthat \divergence \bfu \in L^{p}(\Omega) \}.
\end{gather*}
We are interested in transformations of these Sobolev tensor fields from one domain onto another. 
Suppose that $\Omega, \Omega' \subset \mathbb{R}^3$ are open sets and suppose that $\phi: \Omega \to \Omega'$ is a bi-Lipschitz mapping.
We introduce the gradient-, curl-, and divergence-conforming Piola transformations, respectively, as the mappings 
$\phi^{\grad}: L^{p}(\Omega') \to L^{p}(\Omega)$,
$\phi^{\curl}: \bfL^{p}(\Omega') \to \bfL^{p}(\Omega)$, and 
$\phi^{\divergence}: \bfL^{p}(\Omega') \to \bfL^{p}(\Omega)$.
We also introduce 
$\phi^{\rm b}: L^{p}(\Omega') \to L^{p}(\Omega)$. 
We define them 
for any $v \in L^{p}(\Omega')$ and $\bfw \in \bfL^{p}(\Omega')$ by setting 
\begin{subequations}\label{eq_definition_piola}
\begin{align*}
    \phi^{\grad}(v) &= v \circ \phi, \\
    \phi^{\curl}(\bfw) &= \Jacobian\phi^{T} (\bfw \circ \phi), \\
    \phi^{\divergence}(\bfw) &= \adj(\Jacobian\phi) \left( \bfw \circ \phi \right), \\  
    \phi^{\rm b}(v) &= \det(\Jacobian\phi) \left( v \circ \phi \right),
\end{align*}
\end{subequations}
Here, $\Jacobian\phi$ is the Jacobian matrix of $\phi$ (see also~\cite[Definition~9.8]{ern2021finite}), and $\adj\Jacobian\phi$ denotes taking its adjugate matrix.
These transformations are invertible. 
Bounds on the Lebesgue norms will follow from a more general result below. 
We use the commutativity relations 
\begin{subequations}\label{eq_piola_commute}
\begin{align}
    \grad \phi^{\grad} (v) &= \phi^{\curl}(\grad v), 
    \\
    \curl \phi^{\curl}(\bfv) &= \phi^{\divergence}(\curl \bfv), 
    \\
    \divergence \phi^{\divergence}(\bfw) &= \phi^{\rm b}(\divergence \bfw),
\end{align}
\end{subequations}
where $v \in W^{p}(\grad,\Omega)$, $\bfv \in \bfW^{p}(\curl,\Omega)$, and $\bfw \in \bfW^{p}(\divergence,\Omega)$.  
We summarize this as a commuting diagram:
\begin{align*}
    \begin{CD}
        W^{p}(\grad,\Omega') @>{\grad}>> \bfW^{p}(\curl,\Omega') @>{\curl}>> \bfW^{p}(\divergence,\Omega') @>{\divergence}>> L^{p}(\Omega')
        \\
        @VV{\phi^{\grad}}V 
        @VV{\phi^{\curl}}V 
        @VV{\phi^{\divergence}}V 
        @VV{\phi^{\rm b}}V 
        \\
        W^{p}(\grad,\Omega ) @>{\grad}>> \bfW^{p}(\curl,\Omega ) @>{\curl}>> \bfW^{p}(\divergence,\Omega ) @>{\divergence}>> L^{p}(\Omega ).
    \end{CD}
\end{align*}

@@

We now move the discussion to exterior calculus, beginning with exterior algebra.
Let $V$ be a real vector space. 
Given an integer $k \geq 0$, we let $\Alt^{k}(V)$ denote the space of scalar-valued antisymmetric $k$-linear forms over $V$. 
Recall that any $k$-linear scalar-valued form $u$ over $V$ is called antisymmetric
if 
\begin{gather*} 
    u( v_{\pi(1)}, v_{\pi(2)}, \ldots, v_{\pi(k)} ) 
    = 
    \text{sign}(\pi) 
    u( v_1, v_2, \ldots, v_k ) 
\end{gather*}
for any $v_1, v_2, \dots, v_k \in V$ and any permutation $\pi$ of the indices \(\{1, 2, \ldots, k\}\). 
By definition, $\Alt^{1}(V)$ is just the dual space of $V$, and $\Alt^{0}(V)$ is the space of real numbers. 
Formally, we define $\Alt^{k}(V)$ to be the zero vector space when $k < 0$.

@@

\begin{example}\label{example:shell_simpl}
    Any simplex $T$ (trivially) has a shelling, consisting only of $T$ itself. 
    The boundary complex $\partial \calT(T)$ has a shelling:
    any enumeration of the boundary faces of $T$ constitutes a shelling; see Example 8.2.(iii) in~\cite{ziegler1995lectures}.
\end{example}
\begin{example}
    The standard triangulation of the $3$-dimensional cube by six tetrahedra, the Kuhn triangulation~\cite{kuhn1960some}, is shellable, as are its higher-dimensional generalizations.\footnote{We remark that Kuhn attributes this triangulation to Lefschetz~\cite{lefschetz2015introduction}.}
\end{example}
\begin{example}
    There exists a non-shellable triangulation of a tetrahedron and of a cube in $n=3$, see~\cite[Example 8.9]{ziegler1995lectures}.
\end{example}

@@

        % TODO Check that $\calS'$ is a triangulation
